Get freqs for NWR YD project

### 2020-09-09 getting_frequency_corpus.Rmd
added frequencies from corpus

correspondances=matrix( #list correspondances always by pairs, orthography then phonology
  c(
    #nasal
    ":ii",":ii ",
    ":ee",":ee ",
    ":aa",":aa ",
    ":êê",":êê ",
    ":ââ",":ââ ",
    ":uu",":uu ",
    ":oo",":oo ",

    ":i",":i ",
    ":e",":e ",
    ":a",":a ",
    ":ê",":ê ",
    ":â",":â ",
    ":u",":u ",
    ":o",":o ",

    #non-nasal
    "îî","îî ",
    "êê","êê ",
    "ââ","ââ ",
    "éé","éé ",
    "áá","áá ",
    "óó","óó ",

    "aa","aa ",
    "ee","ee ",
    "ii","ii ",
    "oo","oo ",
    "uu","uu ",

    "î","î ",
    "ê","ê ",
    "â","â ",
    "é","é ",
    "á","á ",
    "ó","ó ",

    "a","a ",
    "e","e ",
    "i","i ",
    "o","o ",
    "u","u ",
    "ú","ú ",
    #consonants from rossel ortho replacements, missing previously
    "mbyw","x ",
    "mbwy","x ",
    "pwy","x ",
    "pyw","x ",
    "tpy","x ",
    "dpy","x ",
    "kpy","x ",
    "myw","x ",
    "mwy","x ",
    "ngw","x ",
    "nmy","x ",
    "ngm","x ",
    "mby","x ",
    "mbw","x ",
    "nty","x ",
    "ndy","x ",
    "nkw","x ",
    "mty","x ",
    "mdy","x ",
    "mgw","x ",
    "dny","x ",
    "dmy","x ",
    "knw","x ",
    "py","x ",
    "pw","x ",
    "ty","x ",
   # "ch"
    "dy","x ",
    "ky","x ",
    "kw","x ",
    #"tp"
    #"dp"
    #"kp"
    "my","x ",
    "mw","x ",
    "ny","x ",
    #"ng"
    "nm","x ",
    #"mb"
    #"nt"
    "nj","x ",
    #"nd"
    #"nk"
    #"mt"
    #"md"
    #"mg"
    #"dn"
    #"dm"
   # "kn"
    #"km"
    #"vy"
    "ly","x ",
    #"lv"
    #"gh"
    #consonants, veeeery dirty!
    "'n","ń ",
    "ń","ń ",
    "tp","tp ",
    "dp","dp ",
    "kp","kp ",
    "ngm","ngm ",
    "ńm","ńm ",
    "ng","ng ",
    "mb","mb ",
    "nd","nd ",
    "nt","nt ",
    "nk","nk ",
    "dn","dn ",
    "kn","kn ",
    "vy","vy ",
    "mt","mt ",
    "md","md ",
    "mg","mg ",
    "dm","dm ",
    "km","km ",
    "gh","gh ",
    "lv","lv ",
    "cch","ch ",
    "ch","ch ", #NOTE
   "w","w ",
   "r","r ",
   "t","t ",
   "y","y ",
   "p","p ",
   "s","s ",
   "d","d ",
  "f","f ",
   "g","g ",
   "h","h ",
   "j","j ",
   "k","k ",
   "l","l ",
   "z","z ",
   "c","c ",
   "v","v ",
   "b","b ",
   "n","n ",
   "m","m "

@@

write.table(wds_uni,"phonemized-words.txt",col.names = F,row.names = F,quote=F)
```

Finally this line happens in the nwr19yd folder

```{bash, eval=F}
tr ' ' '\n' < phonemized-words.txt | sort | uniq -c | sort > phone-counts.txt
```

The final steps were done by hand in an excel file called freq_corpus.xlsx, to calculate frequencies removing items that we think should not count.

>>>>

correspondances=matrix( #list correspondances always by pairs, orthography then phonology
  c(
    #nasal
    ":ii","1 ",
    ":ee","2 ",
    ":aa","3 ",
    ":êê","4 ",
    ":ââ","5 ",
    ":uu","6 ",
    ":oo","7 ",
    ":óó","37 ",

    ":i","8 ",
    ":e","9 ",
    ":a","0 ",
    ":ê","10 ",
    ":â","11 ",
    ":u","12 ",
    ":o","13 ",
     ":ó","38 ",

    #non-nasal
    "îî","14 ",
    "êê","15 ",
    "ââ","16 ",
    "éé","17 ",
    "áá","18 ",
    "óó","19 ",

    "aa","20 ",
    "ee","21 ",
    "ii","22 ",
    "oo","23 ",
    "uu","24 ",

    "î","25 ",
    "ê","26 ",
    "â","27 ",
    "é","28 ",
    "á","29 ",
    "ó","30 ",

    "a","31 ",
    "e","32 ",
    "i","33 ",
    "o","34 ",
    "u","35 ",
    "ú","36 ",
    #consonants from rossel ortho replacements, missing previously
    "mbyw","39 ",
    "mbwy","40 ",
    "pwy","41 ",
    "pyw","42 ",
    "tpy","43 ",
    "dpy","44 ",
    "kpy","45 ",
    "myw","46 ",
    "mwy","47 ",
    "ngw","48 ",
    "nmy","49 ",
    "ngm","50 ",
    "mby","51 ",
    "mbw","52 ",
    "nty","53 ",
    "ndy","54 ",
    "nkw","55 ",
    "mty","56 ",
    "mdy","57 ",
    "mgw","58 ",
    "dny","59 ",
    "dmy","60 ",
    "knw","61 ",
    "py","62 ",
    "pw","63 ",
    "ty","64 ",
   # "ch"
    "dy","65 ",
    "ky","66 ",
    "kw","67 ",
    #"tp"
    #"dp"
    #"kp"
    "my","68 ",
    "mw","69 ",
    "ny","70 ",
    #"ng"
    "nm","71 ",
    #"mb"
    #"nt"
    "nj","72 ",
    #"nd"
    #"nk"
    #"mt"
    #"md"
    #"mg"
    #"dn"
    #"dm"
   # "kn"
    #"km"
    #"vy"
    "ly","73 ",
    #"lv"
    #"gh"
    #consonants, veeeery dirty!
    "tp","Tp ",
    "dp","Dp ",
    "kp","Kp ",
    "ngm","Ngm ",
    "ńm","ñ ",
    "ng","Ng ",
    "mb","Mb ",
    "nd","Nd ",
    "nt","Nt ",
    "nk","Nk ",
    "dn","Dn ",
    "kn","Kn ",
    "vy","Vy ",
    "mt","Mt ",
    "md","Md ",
    "mg","Mg ",
    "dm","Dm ",
    "km","Km ",
    "gh","Gh ",
    "lv","Lv ",
    "cch","CCh ",
    "'n","'n ",
    "ń","ń ",
    "ch","Ch ", #NOTE
   "w","w ",
   "ẃ","ẃ ",
   "r","r ",
   "t","t ",
   "y","y ",
   "p","p ",
   "s","s ",
   "d","d ",
  "f","f ",
   "g","g ",
   "h","h ",
   "j","j ",
   "k","k ",
   "l","l ",
   "z","z ",
   "c","c ",
   "v","v ",
   "b","b ",
   "n","n ",
   "m","m "

@@

wds_uni=unlist(strsplit(wds_uni,split=" "))
counts=data.frame(table(wds_uni))
counts$wds_uni=as.character(counts$wds_uni)

colnames(correspondances)<-c("ortho","fake")
correspondances=data.frame(correspondances)
correspondances$fake=gsub(" ","",as.character(correspondances$fake))

#backtranslate
merge(counts,correspondances,by.x="wds_uni",by.y="fake")->counts
counts=counts[order(counts$Freq),]

colnames(counts)[2]<-"counts_corpus"
counts$freq_corpus=counts$counts_corpus/sum(counts$counts_corpus)

write.table(counts,"segment-counts.txt",col.names = T,row.names = F,quote=T,sep="\t")
```

## Analyses

I ran the code in wrangling again, to integrate these counts & freqs into the main data set.

In addition, I wrote out a new version of segments which has the corpus frequency

```{r}
read.table("segments_with_cor_freq.txt",header=T)->segments # not ran
plot(segments$pc~segments$freq_corpus,type="n")
text(segments$pc~segments$freq_corpus,labels=segments$ortho)

plot(segments$pc~log(segments$freq_corpus),type="n")
text(segments$pc~log(segments$freq_corpus),labels=segments$ortho)

#some stimuli sounds do not occur in the corpus at all
#so we'll give them a really small frequency, just so that they show up
#and we tag them in red
segments$freq_corpus[is.na(segments$freq_corpus)]<-.0001
plot(segments$pc~log(segments$freq_corpus),type="n",main="blue fitted to sounds in corpus, purple to all")
text(segments$pc~log(segments$freq_corpus),labels=segments$ortho,col=ifelse(segments$freq_corpus<.00057,"red","blue"))
#abline(lm(segments$pc~log(segments$freq_corpus),subset=c(segments$freq_corpus<.00057)),col="red",lty=2) #this one is impossible, bc it's vertical, all the x are the same
abline(lm(segments$pc~log(segments$freq_corpus),subset=c(segments$freq_corpus>=.00057)),col="blue",lty=2)
abline(lm(segments$pc~log(segments$freq_corpus)),col="purple",lty=2)
